Algebraic Groups\vspace{-1ex}

### 2019-10-11 Algebraic Geometry/Algebraic-Geometry.tex
Update

>>>>

\section{October 11, 2019}
Today we are going to talk some more about varieties. Coming up on the horizon is a discussion of 
the functor-of-points perspective and we'll talk about Yoneda.

\subsection{Questions/Discussion}
A group of students met up on Wednesday (which we skipped for Yom Kippur) and were talking about 
how to prove the statement: ``points in an affine variety are closed.''

The Nullstellensatz gives us that a point $(x_1,\dots,x_n)\in X$ correspinds to the vanishing locus of $(x_i-a_i)$. 
What if we use the definition in the book, though? If $X=\Spec A=\Hom_{\Alg_k}(A,k)$ (where $A$ is a reduced finitely generated 
$k$-algebra). But then the points are the maximal ideals $\frakm\in k[X]=A$, which are exactly the points we want!

\subsection{Back to Varieties}
Recall the definition of a variety: we are considering $(X,\O_X)$ is a space of functions. Then the idea we want is that we want to say 
that this thing is locally affine. But if you consider infinitely many copies of $\A1$ intersecting pairwise, there is 
an obvious cover by (affine) $\A1$'s. This shouldn't be affine. It doesn't embed in an affine space, for example.

So the definition we used (and the one in Kempf) is that we must have a \textit{finite} cover by affine spaces. We wanted to discuss some examples.
\begin{itemize}
	\item Clearly all affine varieties are varieties.
	\item $\P1$ is our first nontriial example. As a space it is the one-point compactification of $\A1$. The functions 
	$\O_X(U)$ are the rational functions in $t$ with poles not in $U$. You can also construct it by taking the morphism 
	\[\Gm\xrightarrow{t\mapsto 1/t}\Gm\]
	and gluing along this morphism to get a copy of $\P1$! We've already shown in Kempf that it is not affine. 
	What happens if we were to pick the identity above?! We get the line with two origins. It's a variety! But notice that 
	(under the Euclidean topology) the space isn't Hausdorff! 
\end{itemize}
\begin{prob}
	If $X$ and $Y$ are affine varieties, is $X\sqcup Y$ affine?
\end{prob}
\begin{prob}
	How can we expresse the non-Hausdorffness of the line/plane with two origins in the Zariski topology?!
\end{prob}

\subsection{Varieties Glue}
We've been throwing things around, but this is important to write down: Start with $U_i, i\in I$, where $I$ is finite 
(although we could drop finiteness if we don't care about the thing being a variety). For all $i,j$, we have open subsets 
$V_{ij}\subseteq U$ such that $V_{ii}=U_i$ and isomorphisms $\varphi_{ij}:V_{ij}\xrightarrow{\sim}V_{ji}$ of varieties such that 
\begin{enumerate}
	\item $\phi_{ii}=\id$
	\item $\varphi_{ij}(V_{ik}\cap V_{ij})=V_{ji}\cap V_{jk}$
	\item $\forall i,j,k,\, \varphi_{jk}\circ\varphi_{ij}=\varphi_{ik}$ on $V_{ij}\cap V_{ik}$
\end{enumerate}
then there exists a unique package $(X,\iota_i:U_i\hookrightarrow X)$ where $XS$ is a variety and each $\iota_i$ is an open embedding.

One idea is you can rephrase this in categorical language as the colimit of a diagram in a category. Hmmm

### 2019-10-11 Algebraic Groups/Algebraic-Groups.tex
Update

>>>>

\section{October 11, 2019}
Today we're doing problems and stuff. Forgot about that.

\subsection{Casey's Presentation}
Let $G$ and $H$ be objects in a category $\calC$ with finite products. Let $h:H\to G$ be a group homomorphism. That is,

Then we get similar diagrams for the identity and inverse maps (they are respected by $h$). 
Then there is a bunch of diagram work. It's too hard to do a diagram without knowing the shape ahead of time.

Oh hey I used Adam's site! 
\begin{center}
	\begin{tikzcd}
		\ast \arrow[rd, "{(e,e,e)}"]               &                                                                                                                                                  &                                                                                        &                                                                        &                                                                           \\
												   & H\times H\times H \arrow[r, "(i_g\circ h)\times h\times h"] \arrow[d, "1\times m_H"']                                                            & G\times G\times G \arrow[d, "1\times m_g"]                                             &                                                                        &                                                                           \\
												   & H\times H \arrow[r, "(i_G\circ h)\times h"]                                                                                                      & G\times G \arrow[rd, "m_G"]                                                            &                                                                        &                                                                           \\
												   &                                                                                                                                                  &                                                                                        & G                                                                      &                                                                           \\
	\end{tikzcd}
\end{center}
\begin{center}
	\begin{tikzcd}												   & \ast \arrow[r, "e"] \arrow[rd, "{(e,e,e)}"] \arrow[rdddd, "he"', bend right=49]                                                                   & H \arrow[rd, "{(h,h)}"]                                                                &                                                                        &                                                                           \\
												   & &H\times H\times H \arrow[d, "(i_g\circ h)\times h\times h"']                                                                                                                                                                         & G\times G \arrow[rd, "\ast\times 1"] \arrow[ld, "{(\ast,1)\times 1}"',swap] &                                                                           \\
												   &                                                                                                                                                  & G\times G\times G \arrow[d, "1\times m"'] \arrow[rd, "m\times 1"]                      &                                                                        & \ast\times G \arrow[ld, "e\times 1"'] \arrow[lldd, "\pi_1", bend left=49] \\
												   &                                                                                                                                                  & G\times G \arrow[d, "m"']                                                              & G\times G \arrow[ld, "m"]                                              &                                                                           \\
												   &                                                                                                                                                  & G                                                                                      &                                                                        &                                                                           \\
	\end{tikzcd}
\end{center}
\begin{center}
	\begin{tikzcd}
												   & \ast \arrow[rd, "he"', dashed,swap] \arrow[r, "\ast", bend left] \arrow[d, "{(e,e,1)}"'] \arrow[ld, "{(e,e,e)}"'] \arrow[dd, "{(e,e)}", bend left=60] & \ast \arrow[r, "e"]                                                                    & G                                                                      &                                                                           \\
		H\times H\times H \arrow[rd, "1\times m"'] & H\times H\times\ast \arrow[d, "1\times \pi_2"'] \arrow[l, "1\times 1\times e"']                                                                  & G \arrow[u, "\ast"'] \arrow[d, "{(1,1)}"'] \arrow[rd, "{(i,1)}", bend left] &                                                                        &                                                                           \\
												   & H\times H \arrow[r, "h\times h"] \arrow[rr, "(ih)\times h", bend right]                                                                          & G\times G \arrow[r, "i\times 1"]                                                       & G\times G \arrow[uu, "m"]                                              &                                                                          
		\end{tikzcd}
\end{center}
\begin{rmk}
	The idea above is we want to show the first diagram commutes. That is captured in the paths of the second diagram which commutes by 
	the axioms of a group object. The third diagram shows a similar commutativity for the unit $e$.
\end{rmk}
